### 2013-04-23 pnsac/apr13/article_ContactInfo.tex
Initial commit

% Template PNSAC newsletter - Contact information
% Language: Latex

\title{Board Members' Contact Information}

\maketitle

\section{PNSAC Executive}
%\label{sec:misc_pnsac_executive}

%\begin{minipage}{\columnwidth}

\address{Richard Lodge\\
Director, President; Fund Raising\\
613-612-4920\\
\email{rslodgeca@gmail.com}}

\address{Bill Tate\\
Director, Vice President; Special Events, Quarterly Meetings \\
613-523-8817\\
\email{billtate@bell.net}}

\address{Bruce Gemmill\\
Director, Project Manager; Membership\\
613-841-7248\\
\email{dbgemmill@rogers.com}}

\address{Garry Dupont\\
Director, Deputy Project Manager\\
\email{gkdupont@magma.ca}}

%\end{minipage}

%\columnbreak

%\begin{minipage}{\columnwidth}

\address{Bruce Grant\\
Director, NStar Chronicle Editor\\
\email{r\_b\_grant@yahoo.ca}}

\columnbreak

\address{Drew Hodge\\
Director, newsletter typesetting, website, social media\\
\email{ldhodge@gmail.com}}

\address{Phil Chrysler\\
Merchandise\\
\email{chrysler937@rogers.com}}

\address{Roger Button, BA. MA. LLB.\\
Corporate Secretary\\
\email{rgrbttn@gmail.com}}

\address{Paul Labranche\\
Treasurer\\
\email{plabranche@andrews.ca}}

%\end{minipage}

\subsection{Newsletter}

\address{Editor: Bruce Grant}

\address{Typesetter: Drew Hodge}

%\end{multicols}

%\vspace{5mm}

%\hspace{-10mm}
%\setlength\fboxrule{2pt}%
%\framebox{%
%\begin{minipage}{1\columnwidth}%
%\begin{center}
%If you would like to contribute an article to the NStar Chronicle,
%please contact Bruce Grant at 
%\par\end{center}

%\begin{center}
%\email{r_b_grant@yahoo.ca} 
%\par\end{center}

%\begin{center}
%Your comments on the contents of this issue are also appreciated.
%\par\end{center}%
%\end{minipage}}

%\begin{multicols}{2}

\address{PNSAC Newsletter\footnote{This newsletter is typeset using \LaTeX.  The style package used for the newsletter (PNSAC.sty) is a modification of GRASSnews.sty belonging to the Geographic Analysis Resources Support System (GRASS). The modification was made possible by kind permission of the Editor-in-Chief of GRASS-News.} email address:\\
  \email{info@projectnorthstar.ca} -- Attention: Editor}\\
\begin{em}Web site:\end{em} {\normalfont\color{blue}\texttt{\url{www.projectnorthstar.ca}}}\\
%% \begin{em}Facebook page:\end{em} {\normalfont\color{blue}\texttt{\url{www.facebook.com/PNSAC}}}\\
%% \begin{em}Twitter:\end{em} {\normalfont\texttt{@pnsaccomms}}

\begin{figure}
    \centering
    \mbox{\subfigure{ \href{http://www.facebook.com/PNSAC/}{\includegraphics[scale=0.3]{../images/facebook_128x128.eps}}}\quad
    \subfigure{ \href{http://twitter.com/PnsacComms/}{\includegraphics[scale=0.3]{../images/twitter_128x128.eps}}}}
\end{figure}

%\begin{footnotesize}
%This newsletter is typeset using \LaTeX
%  \end{footnotesize}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: main_document.tex
%%% End: