### 2013-04-28 pnsac/apr13/article_RecentEvents.tex

% Template PNSAC newsletter
% Language: Latex

\title{Special Events}
%\subtitle{Visit to the Canada Aviation and Space Museum}
%\author{Bill Tate}

\maketitle

\section{Project North Star Visit to RCAF 426 Squadron ``Thunderbirds''
  -- Friday Jun 14, 2013'}
\label{sec:trenton}

Project North Star is pleased to announce a tour of 426 Squadron at
RCAF Trenton on Friday, June 14th, 2013.  This visit came up at very
short notice, through the help of Tim Timmins (retired Colonel CAF and
member of PNSAC) and Lieutenant Colonel Damon Perrault, Commanding
Officer of 426 Squadron.

Due to the fantastic demand that was generated for our trip to
Montreal, I would strongly recommend an early booking to avoid
disappointment as seats are limited for this fun day of activities.

Our visit to 426 Squadron, which does all the training for the new
C-130J, will involve a tour of their brand new state of the art
training facilities that includes the two airframes that are used for
technician training, another airframe for training loadmasters, a
procedure trainer, and full motion simulators with the usual caveat
``operational requirements.''

After our visit, we will attend the RCAF Memorial Museum which has
undergone a major upgrade to its display area along with their work
shops.

Afterwards we will have a debriefing at Rumour's Restaurant for
schnitzel before returning to Ottawa.

The cost per person is \$70:00. Please send two cheques including a
non-refundable deposit of \$30:00 and a post- dated cheque in the
amount of \$40:00 dated May 14, 2013.

Please note ``Trenton Trip'' on your cheques in addition to sending me
(Bill Tatae) an e-mail message (\email{pnsac.specialevents@gmail.com})
to confirm your participation.  Please include your full name and date
of birth in your e-mail message (this is a requirement of RCAF
Security) as well as your preferred contact information, email or
telephone.

Active volunteers can bring your cheques to the PNSAC office at the
CASM or alternatively mail to:

\address{PNSAC\\
P.O.Box 44005\\
514 Montreal Road\\
Ottawa, ON\\
K1K 4P8}

\section{Third Annual Golf Day at Loch March -- Wednesday August 14, 2013}
\label{sec:golf}

Our Second planned Special Event is for Wednesday August 14th which is
Third Annual Golf Day at Loch March in Carp Ontario.

The day will start off with breakfast at 09:00 with golf starting
shortly after 10:00.

The club professional, as last year, has graciously given PNSAC two
free passes for 18 holes at Loch March, which will be awarded for the
longest drive (staying in bounds of course) and closest to the
pin. Format will be best ball.

Afterwards on the deck, we will discuss the day's game and the
Vice-President will give the winner a ``free beer'' for their effort.

Please email \email{pnsac.specialevents@gmail.com} to confirm you are coming
and call the starter at Loch March 613-839-5885.

Please visit the web site
{\normalfont\color{blue}\texttt{\url{www.lochmarch.com}}} for more
information such as facilities and dress code.

\section{Montreal Area Control Centre and the Bombardier Aerospace  --
Friday Otober 4, 2013}
\label{sec:montreal}

Our Third planned Special Event, on Friday October 4th, is a trip to
the Montreal Area Control Centre and the Bombardier Aerospace factory
in Dorval.

The Montreal Area Control Centre is a secure facility responsible for
Radar Air Traffic Control in the Quebec region including northern
Labrador up to Baffin Island. This tour will have us divided into
smaller groups due to the requirement for quiet so as not to disturb
the Air Traffic Controllers. For operational considerations this tour
could be cancelled on short notice if there is a declared emergency or
an emergency in progress.

After a lunch break we will travel to the Bombardier factory in Dorval
where the CRJ series jet aircraft are built.

For dinner we will proceed to the Willow Inn in Hudson before
proceeding back to Ottawa.

As this tour is now sold out we can accept bookings on a cancellation
basis only.

To make a cancellation booking please email
\email{pnsac.specialevents@gmail.com} to confirm you are willing to standby
for this trip, and mail a non-refundable cheque payable to PNSAC for
\$60:00 payable to PNSAC and in the cheque memo please put in Montreal
trip. Please note this cheque will only be cashed if space becomes
available and if no space becomes available the cheque will be
destroyed.

Please mail the completed booking sheet and cheques to:

\address{PNSAC\\
P.O.Box 44005\\
514 Montreal Road\\
Ottawa, ON\\
K1K 4P8}

\begin{footnotesize}
    \raggedleft PNSAC\\
\end{footnotesize}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: main_document.tex
%%% End: