### 2013-07-14 pnsac/jul13/article_Special.tex

% Template PNSAC newsletter - Article
% Language: Latex
%

% Head

\title{Special Presentation}
%% \author{Drew Hodge}

\maketitle

\end{multicols}

\begin{figure}[ht!]
   \vspace{2em}
   \centering
   %name of the graphic, without the path AND in EPS format:
   \includegraphics[scale=1.0]{roger_button.eps}
   %caption of the figure 
   \caption*{\small \em "Do I really have to wear it?"}
   %label of the figure, which has to correspond to \ref{}:
   \label{fig:rbutton}
\end{figure}

At a Board meeting on March 20, the directors presented a PNSAC golf
shirt to our hard-working Secretary, Roger Button. Our Vice President
Bill Tate made the presentation; the joy and gratitude are evident on
Roger's face.

\begin{footnotesize}
    \raggedleft PNSAC\\
\end{footnotesize}

\begin{multicols}{2}

% End of text.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: main_document.tex
%%% End: