### 2013-07-15 pnsac/jul13/article_ForthEvents.tex

% Template PNSAC newsletter - Miscellany
% Language: Latex

\title{Calendar of Events}

\maketitle

\end{multicols}

\noindent {\bf Note}: The date of the Annual General meeting has been
changed and will now he held on Saturday, June 8, 2013.

\begin{tabbing}
%May 30, 2013     \hspace{20mm}\= Board of Directors' Meeting\\
% {\normalfont\color{red}June 8, 2013}     \>
% {\normalfont\color{red}Annual General Meeting}\\
%June 8, 2013      \> Board of Directors' Meeting (post AGM)\\
%June 27, 2013     \> Board of Directors' Meeting (first meeting of new Board)\\
September 12, 2013     \hspace{20mm}\= Board of Directors' Meeting (to be confirmed)\\
September 21, 2013     \> Members' Quarterly Meeting (to be confirmed)\\
November 28, 2013      \> Board of Directors' Meeting (to be confirmed)\\
December 7, 2013       \> Members' Quarterly Meeting (to be confirmed)\\
March 20, 2014         \> Board of Directors' Meeting (to be confirmed)\\
March 29, 2014         \> Members' Quarterly Meeting (to be confirmed)

\end{tabbing}

\begin{multicols}{2}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: main_document.tex
%%% End: