### 2013-07-15 pnsac/jul13/article_SpecialEvents.tex

% Template PNSAC newsletter
% Language: Latex

\title{Special Events}
%\subtitle{Visit to the Canada Aviation and Space Museum}
%\author{Bill Tate}

\maketitle

\section{Third Annual Golf Day at Loch March -- Wednesday August 14, 2013}
\label{sec:golf}

\begin{figure}[htbp]
   \vspace{2em}
   \centering
   %name of the graphic, without the path AND in EPS format:
   \includegraphics[scale=1.5]{loch_march.eps}
   %caption of the figure 
   %\caption*{\small \em North Star engine 3.}
   %label of the figure, which has to correspond to \ref{}:
   \label{fig:logo}
\end{figure}

A two week break of the restoration of the aircraft will occur in
August, as Mike Irvin takes a well-deserved holiday.

On Wednesday, August 14th, PNSAC will have its third annual golf day
at the beautiful Loch March golf course starting at 10:00 A.M. with a
nine minute separation between flights.  The format for the 18 holes
will be a best ball game.

The club house at Loch March has an excellent breakfast menu so to
make your day more enjoyable it is recommended we meet for breakfast
before the game at 9:00 A.M.

Danielle Nadon, the Club Professional at Loch March has graciously
given us two 18 hole passes for the winner of longest drive and
closest to the pin.

You are encouraged to wear your distinctive PNSAC golf shirt and ball
cap.  As before, guests are welcome.

Please email Danielle at \email{danielle@lochmarch.com} for booking, advising
her that you are part of the Project North Star Day.  Please remember
to send a copy of your e-mail to: \email{pnsac.specialevents@gmail.com}.

Please review the web site\\
{\color{blue}\texttt{\url{http://www.lochmarch.com/about}}} to
familiarize yourself with directions and dress code restrictions to
make your day a ``Great Day at Loch March''.

\begin{itemize}
  \item Seniors mid week rate is \$48:00 and under 60 is \$58:00 plus
    H.S.T
  \item Power cart rental is \$15:49 per person for 18 holes plus
    H.S.T.
  \item Pull cart rental \$5:31 plus H.S.T.
\end{itemize}

\section{Montreal Area Control Centre and the Bombardier Aerospace  --
Friday Otober 4, 2013}
\label{sec:montreal}

Our Third planned Special Event, on Friday October 4th, is a trip to
the Montreal Area Control Centre and the Bombardier Aerospace factory
in Dorval.

The Montreal Area Control Centre is a secure facility responsible for
Radar Air Traffic Control in the Quebec region including northern
Labrador up to Baffin Island. This tour will have us divided into
smaller groups due to the requirement for quiet so as not to disturb
the Air Traffic Controllers. For operational considerations this tour
could be cancelled on short notice if there is a declared emergency or
an emergency in progress.

After a lunch break we will travel to the Bombardier factory in Dorval
where the CRJ series jet aircraft are built.

For dinner we will proceed to the Willow Inn in Hudson before
proceeding back to Ottawa.

As this tour is now sold out we can accept bookings on a cancellation
basis only.

To make a cancellation booking please email
\email{pnsac.specialevents@gmail.com} to confirm you are willing to standby
for this trip, and mail a non-refundable cheque payable to PNSAC for
\$60:00 payable to PNSAC and in the cheque memo please put in Montreal
trip. Please note this cheque will only be cashed if space becomes
available and if no space becomes available the cheque will be
destroyed.

Please mail the completed booking sheet and cheques to:

\address{PNSAC\\
P.O.Box 44005\\
514 Montreal Road\\
Ottawa, ON\\
K1K 4P8}

\begin{footnotesize}
    \raggedleft PNSAC\\
\end{footnotesize}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: main_document.tex
%%% End: