### 2013-10-16 pnsac/newsletter/article_Special.tex
October 2013 newsletter (DRAFT)

% Template PNSAC newsletter - Article
% Language: Latex
%

% Head

\title{PNSAC Merchandise}
%% \author{Drew Hodge}

\maketitle

%\end{multicols}

\begin{figure}[ht!]
   \vspace{2em}
   \centering
   %name of the graphic, without the path AND in EPS format:
   \includegraphics[scale=1.0]{fleece_YOW7578.eps}
   %caption of the figure 
   \caption*{\small \em PNSAC winter apparel}
   %label of the figure, which has to correspond to \ref{}:
   \label{fig:merchandise}
\end{figure}

The first frost warnings on the Weather Channel are reminders that
fall and winter are just around the corner. Keep warm on those frosty
fall days with a PNSAC toque and sweatshirt. We can also take
individual orders for the micro fleece vests shown in the photo on the
left, and if you're looking for early Christmas presents, how about a
coffee mug or a t-shirt?

Have a look at our Facebook page,
{\normalfont\color{blue}\texttt{\href{https://www.facebook.com/media/set/?set=a.610706705637878.1073741851.284829008225651&type=1}{PNSAC
      Merchandise}}}, to see what we're offering, and place your
orders by sending an email message to the following address
\email{pnsac.merchandise@gmail.com}.

For members in the Ottawa-Gatineau region, you can arrange to pick up
your merchandise at the Canada Aviation and Space museum any Monday or
Tuesday, between 10:00 AM and 3:00 PM. We can also ship your purchases
to you for an extra fee. Please specify museum pickup or shipping with
your order.

Merchandise is also available during our Quarterly Meetings
at CAVM.

Look forward to hearing from you.

``The Merchandise Committee''

\begin{footnotesize}
    \raggedleft PNSAC\\
\end{footnotesize}

%\begin{multicols}{2}

% End of text.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: main_document.tex
%%% End: