### 2013-12-14 pnsac/newsletter/article_Feature3.tex

% Template PNSAC newsletter - Article
% Language: Latex
%

% Head

\title{Troop Seats and Extras}
\author{Nelson Smith}

\maketitle

\section{Troop Seats}
\label{troopseats}

Seat back assemblies are completed for the troop seats. Thirty eight
units were produced. The assembling of the backs to the platforms to
produce triple and quad seats will be the next step. This will be done
over the winter.

\begin{figure}[htbp]
   \vspace{2em}
   \centering
   %name of the graphic, without the path AND in EPS format:
   \includegraphics[scale=0.5]{troopseats.eps}
   %caption of the figure 
   \caption*{\small \em New troop seats.}
   %label of the figure, which has to correspond to \ref{}:
   \label{fig:troopseats}
\end{figure}

The prototype is progressing well. As previously stated, measurements
are a rarity with these projects and sometimes it takes longer to
figure out a plan than to do the job. A guesstimate for completion:
early December.

\section{Other Projects}
\label{otherporjects}

Proctective covers for the probes which protrude under the belly of
the North Star were made and installed. The foam pads previously
installed had done their time.

The First Officer's protective seat covers have been manufactured and
installed. This addition was required to shield the seat covers from
UV rays, dust and stains. Similar covers will be made for the
Captain's chair during the winter months.

\begin{footnotesize}
    \raggedleft PNSAC\\
\end{footnotesize}

% End of text.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: main_document.tex
%%% End: