### 2014-05-17 pnsac/newsletter/article_Feature2.tex
Updated files for Apr/May 2014 newsletter.

% Template PNSAC newsletter - Article
% Language: Latex
%

% Head

\title{The Whispering Giant}
\author{Dave Lambeth}

\maketitle

Twinkle, twinkle, old North Star,\\
With wrinkled skin and creaky
spar,\\
For years you’ve always been on
sked,\\
As many of our force you’ve sped\\
to distant lands beyond the skies,\\
you’ve never failed to see the rise\\
of distant suns or moonrise glare.\\
You’ve carried us with tender care\\
-back you came to take us home,\\
from old Japan or France or Rome\\
from England, Goose, and Kef,\\
We cannot hear you.\\
(‘cause we’re deaf)\\

\begin{footnotesize}
    \raggedleft PNSAC\\
\end{footnotesize}

% End of text.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: main_document.tex
%%% End: