### 2014-05-24 pnsac/newsletter/article_Feature3.tex

% Template PNSAC newsletter - Article
% Language: Latex
%

% Head

\title{The Whispering Giant}
\author{Dave Lambeth}

\maketitle

\begin{centering}
Twinkle, twinkle, old North Star,\\
With wrinkled skin and creaky
spar,\\
For years you've always been on
sked,\\
As many of our force you've sped\\
to distant lands beyond the skies,\\
you've never failed to see the rise\\
of distant suns or moonrise glare.\\
You've carried us with tender care\\
-- back you came to take us home,\\
from old Japan or France or Rome\\
from England, Goose, and Kef.\\
We cannot hear you.\\
('cause we're deaf)\\
\end{centering}

\begin{figure}[htbp]
   \vspace{2em}
   \centering
   %name of the graphic, without the path AND in EPS format:
   \includegraphics[scale=0.5]{jim_jung_2.eps}
   %caption of the figure 
   \caption*{\small \em A Whispering Giant (North Star 17515) at
     Resolute Bay NWT in about 1963, taken by Radio Operator Jim Jung.}
   %label of the figure, which has to correspond to \ref{}:
   \label{fig:jim_jung}
\end{figure}

\begin{footnotesize}
    \raggedleft PNSAC\\
\end{footnotesize}

% End of text.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: main_document.tex
%%% End: