### 2016-01-16 pnsac/newsletter/article_Feature2.tex
Adding files for April 2015 newsletter.

% Template PNSAC newsletter - Article % Language: Latex
%

% Head

\title{Volunteers Needed} \author{Richard Lodge}

\maketitle

The Project North Star Association has two types of volunteers. The
first are known as Restoration volunteers and they are the group who
get their hands dirty and physically work on the restoration of the
aeroplane. The other group are known as Association volunteers and
they work in other very important areas of our Association.

At present we are not looking for any more Restoration volunteers but
we do have a need for some Association volunteers. In particular we
are looking for a Social Events Coordinator and a Merchandise Manager
and volunteers who would work with those officers.  These volunteer
posts have become open because the existing volunteers have held the
positions for some years and in each case have decided that it is time
to pursue other interests within the Association.

Below are brief details of the requirements for each volunteer
position. By volunteering for one of these functions, a volunteer will
become more involved with Association activities and will enjoy the
comradeship and rewards of working with a very interesting group of
people who have many diverse life experiences.

The Social Events Coordinator is the person responsible within the
Association for overseeing and organizing events for
members. Typically, he or she would, with the assistance of others,
organize the Members Meetings which are mostly held every three months
and such other events as he or she might agree to run as approved by
the Board of Directors.

The Merchandise Manager is responsible for managing the Association's
stock of merchandise for sale, setting the ordering levels as agreed
with the Board of Directors, placing orders for new stock and ensuring
that sales and inventory levels are properly reported to the
Treasurer. Other volunteers assist the Merchandise Manager in the
work.

If any of our members are interested in taking over one of these
positions or working with these officers or know of a non-member who
would be interested, please would he or she contact me by email at
\email{info@projectnorthstar.ca} or by phone on (613) 612-4920.

\begin{footnotesize} \raggedleft PNSAC\\
\end{footnotesize}

% End of text.

%%% Local Variables: %%% mode: latex %%% TeX-master: main_document.tex
%%% End: