### 2016-01-16 pnsac/newsletter/article_Volunteer.tex

% Template PNSAC newsletter - Article
% Language: Latex
%

% Head

\title{My First Flight}
\author{John Corby, with Bruce Grant}

\maketitle

At the end of the war, the Australians were very anxious to get
British people to immigrate there, and so they had an offering at
\pounds10, almost a free passage. Of course, at the end of the war
there was a distinct lack of shipping, so much had been sunk, and some
of the ships that were left were not really equipped for passenger
carrying anyway.

I put my name down with Shaw Savill Line, and periodically I would go
up to London to Leadenhall Street to see how my name was progressing
down the list. The guy would say "Name? ah yes, Corby", thumb through
the pages and he'd say "About eighteen months---we'll give you a
call". Well, being young and impatient, I didn't like this very much.

The Americans during the war had a vast establishment in Piccadilly
called Rainbow Corner, probably the biggest PX in the world. End of
the war they all went home and the Ontario Government took it over as
an immigration centre.

There was a sign up: \textbf{Emigrate by Air to Ontario Now!
  \pounds70}.  That was a lot more than Australia's \pounds10 of
course, but then it was a lot closer. I thought if I don't like it, I
can come back.  I went in; that would probably be in August of 1947. I
went back for a medical, and on October 13th 1947 I landed in Toronto.

We flew from Heathrow. The aircraft was a North Star, but not in TCA
livery; it was---something else. I think the idea was---they were
trying to prove that they could maintain this schedule across the
Atlantic. They didn't want their name on the planes until they had
proven it. Just in case \ldots

There was at the time a tremendous controversy over these Rolls-Royce
engines. The opposition was all over the government. These engines
were designed for wartime service. In the war, we'd be losing ten to
fifteen planes every day, ninety-five in the Nuremberg raid; that was
the worst day. The engines were designed for maximum performance, not
for reliability and longevity.

When the airlines contracted for these engines, Rolls-Royce guaranteed
that they would do the job. Well, at first they didn't do the job. But
when I flew on them, they obviously did the job!

That was the first time I had ever been in a plane. It took 13 1/2
hours flying time from Heathrow to Toronto, plus a stop in
Gander. They refuelled there and they also fed us a big breakfast---of
pancakes and maple syrup, another first and my introduction to
Canada. When I got to Toronto, my ears rang for two days afterwards
until I got my hearing back.

%\begin{figure}[htbp]
%   \vspace{2em}
%   \centering
%   %name of the graphic, without the path AND in EPS format:
%   \includegraphics[scale=0.5]{jim_jung_2.eps}
%   %caption of the figure 
%   \caption*{\small \em A Whispering Giant (North Star 17515) at
%     Resolute Bay NWT in about 1963, taken by Radio Operator Jim Jung.}
%   %label of the figure, which has to correspond to \ref{}:
%   \label{fig:jim_jung}
%\end{figure}

\begin{footnotesize}
    \raggedleft PNSAC\\
\end{footnotesize}

% End of text.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: main_document.tex
%%% End: