### 2019-01-13 pnsac/newsletter/article_EditorsNotes.tex
January 2019 changes and additions.

% Template PNSAC newsletter - Article % Language: Latex
%

% Head

\title{Editor's Notes} 
\author{Roger Button}

\maketitle

\end{multicols}

\vspace{12pt}

In this issue we are recognizing the 70th anniversary of a very special event
in Canadian aviation history and one which involved a North Star aircraft. The
first non-stop transcontinental flight took place on January 15, 1949 and
readers will find an article dealing with that event in this issue. In
addition, your Association, in partnership with the Canadian Aviation and Space
Museum will be holding a members meeting, and a public viewing of the North
Star on January 19, 2019 to mark this event.

We are also recognizing another one of our long standing members in this
issue. Ted Devey has been active in our organization since the very early
days. We are indeed fortunate that we have a number of other members who have
been with us since the beginning. In 2003 the Ottawa Citizen predicted "a
bright future" for our nascent organization. A copy of the Citizen article is
included in this issue. Not surprisingly, given the challenges involved in
making this project work, the "future" did not always appear as "bright" as
Polaris. However, our ever optimistic President reports that 2018 was
characterized by positive developments in the Project, and our Association's
relationship with CASM.

The Museum's conservator, in the "Conservator's Corner" has contributed a
thoughtful article based on the classic 70's book, \textit{Zen and the Art of
Motorcycle Maintenance} which I am sure
that our readers will enjoy. Good
reading.

%Our Association was incorporated in 2003. We are fortunate that a
%number of the original members of the organization, such as Tim
%Timmins, are still involved as they can provide an historical
%perspective. As Tim says, in the article "Our Members," there have been
%many bumps in the road, and the fact that the aircraft is still a long
%way from completion after 16 years speaks for itself. In January 2017,
%R\'{e}jean (Rej) Demers took up his position as CASM's Conservation Project
%Manager. Rej's first contribution to the Chronicle can be found in the
%"Conservator's Corner." The restoration of the North Star is a priority
%for Rej. He has spent the first year of his mandate gathering
%information and developing systems and procedures which will, in the
%long run, move the restoration of the aircraft forward in a coherent
%manner. I have previously stated the obvious and that is the health and
%vigour of the Association is dependent on the restoration status of the
%aircraft. In the last few years there have been as many bumps in the
%road as there are potholes in Ottawa in the spring. The President is
%very optimistic we have reached a real turning point (see the
%"President's Notes") and wouldn't that be great.

\vspace{12pt}

\begin{footnotesize} \raggedleft PNSAC\\
\end{footnotesize}

\begin{multicols}{2}

% End of text.

%%% Local Variables: %%% mode: latex %%% TeX-master: main_document.tex
%%% End: