### 2019-01-13 pnsac/newsletter/article_PresReport.tex

% Template PNSAC newsletter - Article
% Language: Latex
%

% Head

\title{Notes from the President}
\author{Richard Lodge}

\maketitle

It is always interesting to look back on what was written in previous issues of
any journal. The NStar Chronicle is no exception and so before starting to
write this piece I reread what I had said during the year 2018. For once luck
was with me and as far as I can see I did not write anything that was either
untrue or not achieved during the year.

The last paragraph of my January 2018 Notes suggested that the year would see
increased activity both in the aircraft restoration and within the Association.
Both have proved correct. We have successfully recruited several new
Restoration Volunteers and our Association activities have increased
substantially.

Apart from the occasions when our members either met or have the been actively
taking part in Museum activities we have made a major step forward in that our
Facebook page has again become active under the stewardship of Chris McGuffin.
Chris is a dedicated Restoration Volunteer who can frequently be seen around
the aircraft at a considerable height complete with hardhat and safety harness.
When he returns to the ground, he often gets out his camera and takes
photographs, many of which are now appearing on Facebook. For those of our
readers who are not Facebook subscribers, I encourage you to join Facebook and
to "like" the page Project North Star. You will then see the latest updates on
our restoration progress together with the recent photographs.

Within the next few days our webmaster, Drew Hodge, will be relaunching our
website with an updated homepage and better links to other parts of the
website. Much work has gone into the redesign and we look forward to receiving
comments from users. The website is designed primarily to be a location for all
the Association's historical and archival information. We are planning to add
to the website more PDF versions of previous issues of the NStar Chronicle and
photographs which are not currently available on the site. Current information
will be sent to members by email and updates to our Facebook page.

Our first big event of 2019 will be held at the Museum on Saturday, January
19th to celebrate the 70th anniversary of the first non-stop flight across
Canada between Vancouver and Halifax. The flight was made by a North Star on
15th January 1949. Members have already been advised of the event by email and
on Facebook and I hope to see many of you joining us on that day. Although this
is a Project North Star Association event, we have received considerable help
with planning from the staff of the Museum. This is a good example of the
co-operative working relationship between the Association and Museum.

During the summer we expect to show off the aircraft to the public, as usual,
on Canada Day. We will also be showing the North Star during Doors Open Ottawa
at the beginning of June, if the Museum decides to take part in the event this
year.

During 2018 the Museum organised two very enjoyable events for volunteers.
Invitations were given all Museum registered volunteers including North Star
volunteers. In the summer we once again had a Bush Pilots' Breakfast at which
the Museum staff cook a great breakfast for everyone attending (there is
nothing so good as seeing the Director General, Chris Kitzan flipping
pancakes). Shortly before Christmas the Museum organised a new event, a pot
luck lunch for staff and all the Museum volunteers---another thoroughly good
overeating occasion.

One of the benefits of becoming an active member of PNSAC is for the
comradeship with other members who all have an interest in aviation. It is not
necessary to be a volunteer to enjoy the company of other members. Once a month
there is an informal breakfast held in a caf\'{e} on Arch St which is relatively
near the Ottawa General Hospital. The breakfast started about two years ago
under the leadership of Bill Tate. Each month we chat about aviation, the North
Star project and whatever else we feel like talking about. I usually give the
group an update on anything of interest that has happened since the last
breakfast. Joining this informal group is a good way to become involved with
the Association and to possibly move on to taking a more active role in the
Association.

We will be holding our Annual General Meeting at the Museum on Saturday, April
13. At this meeting we re-elect our Board of Directors and would welcome the
nomination of any member who might be interested in serving on our Board. At
present we have five members. This leaves one vacancy to bring us up to our
normal Board number of six members.

2019 is starting well. We look forward further significant milestones in the
North Star restoration process such as rehanging of the restored Engine \#4. We
are also hoping to arrange further events for our Association members.

\begin{footnotesize}
    \raggedleft PNSAC\\
\end{footnotesize}

% End of text.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: main_document.tex
%%% End: