### 2020-02-24 pnsac/newsletter/article_EditorsNotes.tex
Draft updates for Feb-Mar 2020.

% Template PNSAC newsletter - Article % Language: Latex
%

% Head

\title{Editor's Notes} 
\author{Roger Button}

\maketitle

\end{multicols}

\vspace{12pt}

Putting out this edition of the Chronicle has been an example of great
co-operation from members and non-members alike. Thanks are owed to a
number of people. Our member Gary Whitten was instrumental in getting
the necessary permission from Air Force Magazine to publish the
"Diversion to Alert" article.  Allan Bowes, another member, suggested
that we should try to get the "Civil Merlins" article into an edition
of the Chronicle.  Following his suggestion we contacted Ben Dunnell,
the editor of Aeroplane Magazine. Through Ben's offices we were able to
secure permission to reprint this article.  The "Civil Merlins" was
part of a comprehensive 17 page Aeroplane Database article on the
Merlin. If any members wish to read the entire article it was published
in the July 2017 issue.

In addition thanks go out to the other contributors, namely Richard
Lodge, Garry Dupont,  Neil Raynor and Bruce Gemmill. Bruce took the
lead in putting together the Progress Report.  Finally, I would like to
thank Drew Hodge our stalwart publisher of this edition. We are always
on the lookout for content for the Chronicle so please do not hesitate
to bring forward any ideas that you might have. Enjoy.

\vspace{12pt}

\begin{footnotesize} \raggedleft PNSAC\\
\end{footnotesize}

\begin{multicols}{2}

% End of text.

%%% Local Variables: %%% mode: latex %%% TeX-master: main_document.tex
%%% End: