### 2020-03-08 pnsac/newsletter/article_EditorsNotes.tex
PNSAC newsletter--Mar.2020.

% Template PNSAC newsletter - Article % Language: Latex
%

% Head

\title{Editor's Notes} 
\author{Roger Button}

\maketitle

\end{multicols}

\vspace{12pt}

Putting out this edition of the Chronicle has been an example of great
co-operation from members and non-members alike. Thanks are owed to a
number of people. Our member Gary Whitten was instrumental in getting
the necessary permission from Air Force Magazine to publish the
"Diversion to Alert" article.

%Allan Bowes, another member, suggested
%that we should try to get the "Civil Merlins" article into an edition
%of the Chronicle.  Following his suggestion we contacted Ben Dunnell,
%the editor of Aeroplane Magazine. Through Ben's offices we were able to
%secure permission to reprint this article.  The "Civil Merlins" was
%part of a comprehensive 17 page Aeroplane Database article on the
%Merlin. If any members wish to read the entire article it was published
%in the July 2017 issue.

In addition thanks go out to the other contributors, namely Richard
Lodge, Garry Dupont,  Neil Raynor, and R\'{e}jean Demers. A special thanks
to Bruce Gemmill who took the lead in putting together the progress report,
which can be found in the Conservator's Corner section.

Finally, I would like to thank Drew Hodge our stalwart publisher of this
edition. We do appreciate receiving feedback from our readers including
comments on the articles or suggestions for new articles. We are always on the
lookout for content for the Chronicle so please do not hesitate to bring
forward any ideas that you might have. As noted in the Conservator's Corner
article the restoration of engine number 4 has been completed. Our next
edition will focus on the engine restoration work so if you have any material
or comments specific to that aspect of the restoration please let us know.
Enjoy.

\vspace{12pt}

\begin{footnotesize} \raggedleft PNSAC\\
\end{footnotesize}

\begin{multicols}{2}

% End of text.

%%% Local Variables: %%% mode: latex %%% TeX-master: main_document.tex
%%% End: