### 2020-09-26 pnsac/newsletter/article_EditorsNotes.tex
Sep2020--first draft.

% Template PNSAC newsletter - Article % Language: Latex
%

% Head

\title{Editor's Notes} 
\author{Roger Button}

\maketitle

\end{multicols}

\vspace{12pt}

In this issue we are recognizing the completion of a very important step in the
restoration of the North Star, namely, its Merlin engines, and our members who
have volunteered with CASM to carry out the necessary work.  This is a very
informative issue and I am sure we will all learn a great deal. I was not
aware, until putting this edition together, that the Merlin designation was in
keeping with the custom of Rolls Royce to name its aircraft engines after birds
of prey.

The article, "Four Merlins" provides the reader with a detailed history of the
restoration of these engines.   Due to its length, the article will be divided
into two editions;  starting with engines number one and two in this edition;
and three and four in the next edition. We are fortunate to have members who
have had  lengthy involvement with this part of the project. Sadly one of those
members passed away recently. Inside you will find a tribute to Tim Timmins by
the President of our Association.

We look forward to receiving your comments and suggestions regarding this
edition and for future editions whether good, bad or ugly. Please use my email
address located on the back page.

Good reading.

\vspace{12pt}

\begin{footnotesize} \raggedleft PNSAC\\
\end{footnotesize}

\begin{multicols}{2}

% End of text.

%%% Local Variables: %%% mode: latex %%% TeX-master: main_document.tex
%%% End: