### 2020-09-26 pnsac/newsletter/article_Feature2.tex

% Template PNSAC newsletter - Article % Language: Latex
%

% Head

\title{The Civil Merlin} 
\author{Dr. Jakob Whitfield}
\maketitle

\textit{The following article is an extract from a larger piece which appeared
in the July 2017 edition of Aeroplane magazine under the heading
"Rolls-Royce Merlin."  We wish to thank Allan Bowes for bringing it to
our attention. This article is republished with the kind permission of
the publishers of the magazine and the author Dr. Jakob Whitfield.}\\

The earliest Merlins to operate in a civil mode were the Merlin T24 series,
developed in 1944. These were single-stage twin-speed units similar to the
Merlin 24s fitted to the RAF's Lancasters, but were modified to improve service
life, and were fitted to Transport Command's Avro Yorks. Long-range transport
operation entailed running at relatively low cruise powers for long
periods---under these conditions the lower cylinder head temperatures caused
deposits of
lead oxide from the fuel, resulting in excessive spark plug fouling. To counter
this, the Merlin T24/4 incorporated a charge heater to increase the inlet
temperature. Post-war, the Merlin 500 series essentially comprised civil and
export versions of the T24, incorporating its modifications, though only the
Merlin 501 included a charge heater.

The 600 and 700-series engines were two-speed two-stage engines, based on the
military 100-series. They were fitted with the so-called 'transport heads and
banks', strengthened for greater reliability. Most marks had some form of
variable intercooling to allow for charge heating to reduce plug fouling under
cruise conditions. Initially the system was plagued by coolant leaks, but as
experience showed that zero intercooling at cruise allowed enough charge
heating to reduce leading, a simple stop valve was fitted to the system. This
allowed full intercooling for take-off and climb, which could then be turned
off for cruise.

Civil Merlins were fitted to Avro Lancastrians, Yorks and Tudors, but their
flagship role was on the Canadair DC-4M North Star, a Douglas DC-4-derived
design intended for Trans-Canada Air Lines (TCA). The North Stars used a
modified Universal Power Plant installation, as the DC-4's nacelle bulkheads
were slightly larger than the SBAC standard. In practice it turned out the
Merlins were not ideally suited for use in civil operations: though the North
Stars flew higher and faster than DC-4s, their time between engine overhauls
was around 850 hours, roughly a third of that of commercial US radial engines.
The engines also produced a lot of cabin noise, though this was alleviated
somewhat by revised 'cross-over' exhausts that ducted the cabin-side stacks
outboard, developed first by TCA and then by Rolls-Royce themselves.

The extra running costs were mostly borne by Rolls-Royce. When TCA's managing
director expressed dissatisfaction with the Merlin's commercial performance,
Hives asked what a reasonable level of maintenance cost would be. Being told
\$4 per engine hour, he agreed to service the fleet's engines for this amount.
This early form of 'power by the hour' was initially expensive for Rolls-Royce,
but by the end of the Merlin's life the company had learned enough to
supposedly make a small profit at this level.

What was undoubtedly true was that Rolls-Royce learned a great deal about the
harsh realities of commercial operation in a short time. Hives is supposed to
have said to TCA, "We didn't know the Merlin until you started operating it!"
In the longer term, this paid off, TCA---later Air Canada---selecting
Rolls-Royce engines for its future fleet: Darts in Viscounts, Tynes in
Vanguards, Conways in DC-8s, and RB211s in TriStars.

\begin{footnotesize}
  \raggedleft PNSAC\\
\end{footnotesize}

% End of text.

%%% Local Variables: %%% mode: latex %%% TeX-master: main_document.tex
%%% End: