### 2020-09-26 pnsac/newsletter/article_Special.tex

% Template PNSAC newsletter - Article
% Language: Latex
%

% Head

\title{Remembering Austin (Tim) Timmins 1928-2020}
\author{Richard Lodge}

\maketitle

\begin{figure}[htbp]
   \vspace{2em}
   \centering
   %name of the graphic, without the path AND in EPS format:
   \includegraphics[scale=0.75]{tim-timmins.png}
   %caption of the figure 
   \caption*{\small \em Austin (Tim) Timmins.}
   %label of the figure, which has to correspond to \ref{}:
   \label{fig:tim}
\end{figure}

We were all very sorry to hear that our second President and also Life Member,
Tim Timmins, had passed away on September 7th. Tim was a founding member of the
Project North Star Association of Canada and one of its original directors. He
was an active member of the small group of aviation enthusiasts who worked for
many months with the Canada Aviation \& Space Museum to negotiate a
groundbreaking arrangement to restore Canada's sole remaining North Star
aircraft to a standard suitable for display in the Museum.

Tim became a dirty hands volunteer in the early years of the project and was
involved in the often difficult and usually very dirty initial restoration work
on the aircraft. He will have taken part in "bird nesting" when volunteers
spent many hours removing years of bird nest debris from various parts of the
aircraft.

Following the untimely death of our first President, Robert Holmgren in October
2007, Tim agreed to become our second President. He held this position until
June 2000 during which time he was instrumental in helping the Association move
from a fledgling operation to a respected part of the Aviation Museum's
volunteer team.

Many of us will remember Tim's cheery smile and the long hours he spent talking
to members of the public during the Museum's open days and public events. He
always had anecdotes and stories to tell of his time in the RCAF as a navigator
on the North Star.

\begin{footnotesize}
    \raggedleft PNSAC\\
\end{footnotesize}

% End of text.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: main_document.tex
%%% End: