### 2021-08-13 pnsac/newsletter/article_EditorsNotes.tex
Changes for issue 17-1

% Template PNSAC newsletter - Article % Language: Latex
%

% Head

\title{Editor's Notes} 
\author{Roger Button}

\maketitle

\end{multicols}

\vspace{12pt}

Unfortunately COVID-19 and its restrictions continue to hamper the restoration
of the North Star. So it's a good time to look at what has been accomplished to
date by the restoration team which consists of both Museum staff and volunteers.
This edition will continue the focus on the Merlin engines. As many of you know
many of the volunteers are also members of our organization.

You will find photos of some of those volunteers in the technical article on
engines three and four. You will also find a interview with Garry Dupont who has
had a leading role in the engine shop work.

In the next edition we expect to have more articles about our members
experiences in the restoration project.

In the meantime if you have any comments on this edition or suggestions for
future editions please send them to me.

\vspace{12pt}

\begin{footnotesize} \raggedleft PNSAC\\
\end{footnotesize}

\begin{multicols}{2}

% End of text.

%%% Local Variables: %%% mode: latex %%% TeX-master: main_document.tex
%%% End: