### 2022-05-20 pnsac/newsletter/article_EditorsNotes.tex
Updated for final May 2022 issue.

% Template PNSAC newsletter - Article % Language: Latex
%

% Head

\title{Editor's Notes} 
\author{Roger Button}

\maketitle

\end{multicols}

\vspace{12pt}

This edition marks a change of leadership of our association. Richard Lodge, our longtime president, retired at the end of 2021. His replacement is Chris McGuffin. Richard's reminiscences of his time as president can be found in the Our Members column.  While Richard's story talks about his past experiences, Chris' Notes from the President sets out his vision for the future of the organization. R\'{e}jean Demers, CASM's Project Manger for the North Star also addresses, in the Conservator's Corner column,  how he envisages the Project restarting this year.  Finally, there is a fascinating article by John Makadi and Christ McGuffin about how they found and acquired an APS 42 radar unit for the aircraft. Good reading.

\vspace{12pt}

\noindent As always comments and suggestions for future editions are welcomed.

\begin{footnotesize} \raggedleft PNSAC\\
\end{footnotesize}

\vspace{60pt}

\begin{multicols}{2}

% End of text.

%%% Local Variables: %%% mode: latex %%% TeX-master: main_document.tex
%%% End: